---
title: "20241127_p8105_hw6"
output: github_document
date: "2024-11-28"
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(rnoaa)
library(broom)
library(purrr)
library(tidyverse)
```

## Problem 1

```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2017-01-01",
    date_max = "2017-12-31") %>%
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) %>%
  select(name, id, everything())
```

Steps: 

1. Generate 5000 bootstrap samples 

2. Fit a linear regression model for each bootstrap sample where response is tmax and predictor is tmin

3. For each bootstrap sample, calculate the R^2 and the logarithm of estimated coefficients multiplied

Boostrap function for 5000 samples
```{r}
set.seed(123)

bootstrap_results = replicate(5000, {

sample_indices = sample(seq_len(nrow(weather_df)), replace = TRUE)
bootstrap_sample = weather_df[sample_indices, ]
  
  model = lm(tmax ~ tmin, data = bootstrap_sample)
  model_summary = summary(model)
  
  b0 = coef(model)[1]
  b1 = coef(model)[2]
  
  r_squared = model_summary$r.squared
  
  log_b0_b1 = log(b0 * b1)

  return(c(r_squared = r_squared, log_b0_b1 = log_b0_b1))
}, 

simplify = "matrix")

bootstrap_df = as.data.frame(t(bootstrap_results))
colnames(bootstrap_df) = c("R_squared", "log_b0_b1")

```


Plotting my results

```{r}
ggplot(bootstrap_df, aes(x = R_squared)) +
  geom_histogram(binwidth = 0.01, fill = "blue", alpha = 0.7, color = "black") +
  labs(title = "Bootstrap Distribution of R-squared",
       x = "R-squared",
       y = "Frequency") +
  theme_minimal()
```

The histogram plot for the R^2 values resulting from the 5000 bootstrap shows a unimodal distribution, centered around the mean of the bootstrap R^2 values. Additionally, the mean of the bootstrap R^2 values should be very close to the R^2 value of the original weather dataframe that we bootstrapped 5000 samples from. Most R^2 values lie within a range indicating a good fit of the linear model. The variability is due to the resampling process of bootstrapping. 


```{r}
ggplot(bootstrap_df, aes(x = log_b0_b1)) +
  geom_histogram(binwidth = 0.1, fill = "green", alpha = 0.7, color = "black") +
  labs(title = "Bootstrap Distribution of log(b0^ * b1^)",
       x = "log(b0^ * b1^)",
       y = "Frequency") +
  theme_minimal()
```



